---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# bdsreader

<!-- badges: start -->
[![R-CMD-check](https://github.com/growthcharts/bdsreader/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/growthcharts/bdsreader/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

The `bdsreader` package is a lightweight package that

- Defines JSON data schema's for electronic data exchange of child-level data based on the Basisdataset JGZ 4.0.1;
- Reads and parses child data coded according to the JSON data schema;
- Compares the child data against the JSON validation schema;
- Calculates the child's D-score from the Van Wiechen Schema (DDI) responses;
- Adds Z-scores for height, weight, head circumference, BMI, weight-for-height and D-score;
- Converts the result into a a list with person-level and time-level data for use in JAMES.

The `bdsreader` translates child data (incoming via an API request) into a data object useful for `R` processing. The package is part of Joint Automatic Measurement and Evaluation System (JAMES) developed by the Netherlands Organisation for Applied Scientific Research TNO.

## Installation

Install the development version `bdsreader` by

```{r eval = FALSE}
install.packages("remotes")
remotes::install_github("growthcharts/bdsreader")
```

There is no CRAN release.

## Example

The following commands illustrate the main use of `bdsreader`.

```{r}
library(bdsreader)
fn <- system.file("examples", "maria.json", package = "bdsreader")
tgt <- read_bds(fn)
timedata(tgt)
```

Column `age` holds decimal age for the measurement. Every row contains a measurement `yname`, the conditioning variable `xname` and the Z-score `zname`. The column named `zref` holds the name of the growth reference (as defined in the `nlreference` package) used to calculate the Z-score. Columns `y`, `x` and `z` store their values, respectively. 

The `persondata()` function extracts the person-level information:

```{r}
persondata(tgt)
```

The result of `read_bds()` feeds into further data processing in `R`.

## Breakdown in steps

### JSON Input Data

The example file `maria.json` contains Maria's data coded in JSON format according to BDS-schema file [bds_v3.0.json](https://raw.githubusercontent.com/growthcharts/bdsreader/master/inst/schemas/bds_v3.0.json). Here's the contents of the file with the child data:

```{javascript}
{
  "Format": "3.0",
  "organisationCode": 1234,
  "reference": "Maria",
  "clientDetails": [
    {
      "bdsNumber": 19,
      "value": "2"
    },
    {
      "bdsNumber": 20,
      "value": "20181011"
    },
    {
      "bdsNumber": 82,
      "value": 189
    },
    {
      "bdsNumber": 91,
      "value": "1"
    },
    {
      "bdsNumber": 110,
      "value": 990
    },
    {
      "bdsNumber": 238,
      "value": 1670
    },
    {
      "bdsNumber": 240,
      "value": 1900
    }
  ],
  "clientMeasurements": [
    {
      "bdsNumber": 235,
      "values": [
        {
          "date": "20181111",
          "value": 380
        },
        {
          "date": "20181211",
          "value": 435
        }
      ]
    },
    {
      "bdsNumber": 245,
      "values": [
        {
          "date": "20181011",
          "value": 990
        },
        {
          "date": "20181111",
          "value": 1250
        },
        {
          "date": "20181211",
          "value": 2100
        }
      ]
    },
    {
      "bdsNumber": 252,
      "values": [
        {
          "date": "20181111",
          "value": 270
        },
        {
          "date": "20181211",
          "value": 305
        }
      ]
    }
  ],
  "nestedDetails": [
    {
      "nestingBdsNumber": 62,
      "nestingCode": "01",
      "clientDetails": [
        {
          "bdsNumber": 63,
          "value": "19950704"
        }
      ]
    },
    {
      "nestingBdsNumber": 62,
      "nestingCode": "02",
      "clientDetails": [
        {
          "bdsNumber": 63,
          "value": "19901202"
        }
      ]
    }
  ]
}
```

JSON is a lightweight format to exchange data between electronic systems. Field `"bdsNumber"` refers to the numbers defined in the Basisdataset JGZ. Field `"value"` contains the value for the `"bdsNumber"`. See [Basisdataset JGZ](https://www.ncj.nl/onderwerp/digitaal-dossier-jgz/basisdatasetjgz-bds/) 4.0.1 for more details on `"bdsNumber"`.


### Read and parse input data

### Validate input data

### Age calculation

### D-score calculation

### Z-score calculation

### Structure of result

## Further information

