---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# bdsreader

<!-- badges: start -->
[![R-CMD-check](https://github.com/growthcharts/bdsreader/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/growthcharts/bdsreader/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

The `bdsreader` package is a lightweight package that

- Defines JSON data schema's for electronic data exchange of child-level data based on the Basisdataset JGZ 4.0.1;
- Reads and parses child data coded according to the JSON data schema;
- Compares the child data against the JSON validation schema;
- Calculates the child's D-score from the Van Wiechen Schema (DDI) responses;
- Adds Z-scores for height, weight, head circumference, BMI, weight-for-height and D-score;
- Converts the result into a a list with person-level and time-level data for use in JAMES.

The `bdsreader` translates child data (incoming via an API request) into a data object useful for `R` processing. The package is part of Joint Automatic Measurement and Evaluation System (JAMES) developed by the Netherlands Organisation for Applied Scientific Research TNO.

## Installation

Install the `jamesdemodata` and `bdsreader` packages as

```{r eval = FALSE}
install.packages("remotes")
remotes::install_github("growthcharts/jamesdemodata")
remotes::install_github("growthcharts/bdsreader")
```

There is no CRAN release of these packages.

## Examples

`read_bds()` and `write_bds()` are the main functions of the package. 

### Reading JSON child data

We work with an example dataset `maria.json`, containing the data of one child, from the `jamesdemodata` package. Find the filename and show the first 10 lines:

```{r}
fn <- system.file("json", "examples", "maria.json", package = "jamesdemodata")
cat(paste(readLines(fn, n = 10), collapse = "\n"), "\n...")
```

The following commands illustrate the use of the `read_bds()`` function:

```{r}
library(bdsreader)
child <- read_bds(fn)
class(child)
```

`read.bds()` returns a `bdsreader` object. We print it as

```{r}
child
```

The object is a list with two elements: `psn` and `zyx`: 

- The `psn` element contains the person-level information;
- The `xyz` element contains time-varying measurements for the child.

See `?init_bdsreader` for more details on the structure of the `bdsreader` class. 

### Writing JSON child data

The `write_bds()` function perform the inverse operation. By default, it returns minified JSON:

```{r}
js <- write_bds(child)
cat(paste(js, collapse = "\n"), "\n...")
```

You can write the data in a pretty format using the `indent` and to a file using `file` argument.



