---
title: "Steps to read JSON child data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Steps to read JSON child data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, include= FALSE}
library(bdsreader)
knitr::opts_chunk$set(echo = TRUE)
```

The remainder of this README describes the steps taken to read the data in more detail. The steps are:

### JSON Input Data

The example file `maria.json` contains Maria's data coded in JSON format according to BDS-schema file [bds_v3.0.json](https://raw.githubusercontent.com/growthcharts/bdsreader/master/inst/schemas/bds_v3.0.json). Here's the contents of the file with the child data:

```{javascript}
{
  "Format": "3.0",
  "organisationCode": 1234,
  "reference": "Maria",
  "clientDetails": [
    {
      "bdsNumber": 19,
      "value": "2"
    },
    {
      "bdsNumber": 20,
      "value": "20181011"
    },
    {
      "bdsNumber": 82,
      "value": 189
    },
    {
      "bdsNumber": 91,
      "value": "1"
    },
    {
      "bdsNumber": 110,
      "value": 990
    },
    {
      "bdsNumber": 238,
      "value": 1670
    },
    {
      "bdsNumber": 240,
      "value": 1900
    }
  ],
  "clientMeasurements": [
    {
      "bdsNumber": 235,
      "values": [
        {
          "date": "20181111",
          "value": 380
        },
        {
          "date": "20181211",
          "value": 435
        }
      ]
    },
    {
      "bdsNumber": 245,
      "values": [
        {
          "date": "20181011",
          "value": 990
        },
        {
          "date": "20181111",
          "value": 1250
        },
        {
          "date": "20181211",
          "value": 2100
        }
      ]
    },
    {
      "bdsNumber": 252,
      "values": [
        {
          "date": "20181111",
          "value": 270
        },
        {
          "date": "20181211",
          "value": 305
        }
      ]
    }
  ],
  "nestedDetails": [
    {
      "nestingBdsNumber": 62,
      "nestingCode": "01",
      "clientDetails": [
        {
          "bdsNumber": 63,
          "value": "19950704"
        }
      ]
    },
    {
      "nestingBdsNumber": 62,
      "nestingCode": "02",
      "clientDetails": [
        {
          "bdsNumber": 63,
          "value": "19901202"
        }
      ]
    }
  ]
}
```

JSON is a lightweight format to exchange data between electronic systems. Field `"bdsNumber"` refers to the numbers defined in the Basisdataset JGZ. Field `"value"` contains the value for the `"bdsNumber"`. See [Basisdataset JGZ](https://www.ncj.nl/onderwerp/digitaal-dossier-jgz/basisdatasetjgz-bds/) 4.0.1 for more details on `"bdsNumber"`.


### Read and parse input data

### Validate input data

### Age calculation

### D-score calculation

### Z-score calculation

### Structure of result

## Further information

